---
title: "Dual Tech Patent Classification"
author: "ZH"
date: "2026-01"
output: 
  html_document:
    self_contained: true
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(plotly)

set.seed(972)
```

Dual-use scoring of patents based on semantic similarity between patent abstracts and items on the Wassenaar Arrangement (WA) control lists.

## Proof of Concept

### Data:

1) 50,000 US patents granted between 2015–2020 (sourced from lens.org).

2) Wassenaar Arrangement (WA) control list items for dual-use technologies (https://www.wassenaar.org/control-lists/). These lists define controlled categories used by Israeli arm export-control.

### Method:

1) Text representation of both patent abstracts and WA control-list text (item list by category). 

  - Current baseline uses TF-IDF; a better approach is embeddings to better capture semantic similarity.

2) Compute a patent-level 'duality score' using cosine similarity:

  - compute cosine similarity between each patent abstract and each WA text.

  - assign each patent the maximum similarity across WA texts.

Next step: select a classification threshold to flag likely dual-use patents.

### Example:

Sample of 500 patents from the data, reporting for each patent: the duality_score, the best-matching WA item, and the matched text snippet (move mouse across bars).

```{r plot, results = "asis", echo=FALSE}
# read data
patent_data_sample <- 
  readRDS(here("data", "patent_daul_tech_sim_data.RDS")) |> 
  slice_sample(n = 500) # keep sample of 500 patents

# add aux data and prepare for plot
patent_data_sample <- 
  patent_data_sample |> 
  mutate(
    cos_decile = ntile(cosine_sim, 10),
    cos_decile = factor(cos_decile, levels = 1:10),
    patent_title = forcats::fct_reorder(patent_title, cosine_sim),
    tooltip = paste0(
      "Patent: ", patent_title,
      "<br>WA_doc: ", WA_catagory,
      "<br>Owner: ", Owners,
      "<br>Abstract: ", str_wrap(Abstract, width = 50) |> str_replace_all("\n", "<br>"),
      "<br>Cosine: ", round(cosine_sim, 3),
      "<br>Decile: ", cos_decile
    )
  )


# plot
pal_traffic10 <- colorRampPalette(c("red", "yellow", "green"))(10)

p <- 
  ggplot(patent_data_sample, aes(x = patent_title, y = cosine_sim,
                       fill = cos_decile, text = tooltip)) +
  geom_col(position = position_dodge2()) +
  scale_fill_manual(values = pal_traffic10, name = "Cosine decile") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )

ggplotly(p, tooltip = "text")

rm(patent_data_sample, p, pal_traffic10)
```

<!-- https://www.nber.org/system/files/working_papers/w31925/w31925.pdf -->